SiO Molecular Jets around young stars - A numerical perspective

### 2013-10-15 PAPER/Paola_1/ms_ver2.tex
all but one comments from paola in rad trans section addressed

\subsection{The radiative transfer code} \label{subsec:radiative_transfer_code}
The radiative transfer program used is LIME (LIne Modeling Engine; \citealt{Brinch:2010p13078}), which  calculates line intensities based on a weighted sample of randomly chosen points in a continuous 3D model. The method of selecting these points is given in section \ref{subsec:gridding}. At each of these points, the density of the main collision partner (equivalent amount of H$_2$, given by n(H$_2$)+ 0.5 n(H), assuming that atomic Hydrogen has half the SiO collisional cross section of molecular hydrogen and that ionized Hydrogen does not collide with SiO), gas and dust temperatures, velocity, molecular abundances and unresolved turbulent velocity are specified. These points are then smoothed by Lloyd's algorithm \citep{Lloyd1982} in order to minimise the variation in distance between points whilst keeping the same underlying distribution. These points are then connected by Delaunay triangulation and it is between the points connected by this method that photon are allowed to propagate (fig.~\ref{grid}). The level populations of the selected molecules are calculated at each of these points from collisional and radiative (de)excitation and the local radiation field is calculated. This is repeated 20 times with the populations of each level converging towards a single value. This number of iterations is sufficient for the signal to noise ratio of the level populations (as defined in \citealt{Brinch:2010p13078}) to exceed 1000 for 99\% of the points, ensuring that the simulation has converged on a stable level population. After 20 iterations the model is ray-traced in order to produce synthetic brightness maps. The average of ten separate runs was taken to minimise the artefacts in the output images, resulting from the grid construction.% (Fig.~\ref{points}).

@@

\subsection{Grid construction} \label{subsec:gridding}
In order to construct the grid, candidate points are randomly selected from the volume to be simulated. These candidates then have their equivalent H$_2$ number density, and the number density of SiO, compared against those of a reference point in order to decide if the candidate point is to be used in the grid or not. Candidate grid points are selected at random in a cylindrical coordinate system that is linearly spaced in z and $\theta$ and logarithmically spaced in r. For each point to be selected, a random number $\alpha$ is drawn from the semi-open set [0,$\,$1) as a threshold. After selection of random coordinates, the H$_2$ density and SiO density at the candidate point (n and m, respectively) are compared against the densities of a reference point in the un-perturbed ambient medium multiplied by a factor of $\frac{4\eta}{5}$ (n$_0$ and m$_0$). If $\alpha<\left( \frac{n}{n_0} \right)^{0.3}$ or $\alpha< \left( \frac{m}{m_0} \right)^{0.3}$ then the point is selected for use. Otherwise another r, $\theta$, z co-ordinate is selected and it becomes the candidate point. In addition to this method of selection, 5\% of the points are linearly distributed in x, y and z with no bias with regards to density or abundance. This provides a minimum level of sampling for the large low density regions in the outer parts of the simulated volume. %See fig. \ref{points} for an example of the points distribution in r, z. The function comparing the candidate point to the reference point and the candidate point distribution were selected empirically to sample all scales while ensuring that the majority of points are located in the inner disc where the density is higher.  \smallskip

>>>>

\subsection{The radiative transfer code} \label{subsec:radiative_transfer_code}
The radiative transfer program used is LIME (LIne Modeling Engine; \citealt{Brinch:2010p13078}), which  calculates line intensities based on a weighted sample of randomly chosen points in a continuous 3D model. The method of selecting these points is given in section \ref{subsec:gridding}. At each of these points, the density of the main collision partner (equivalent amount of H$_2$, given by n(H$_2$)+ 0.5 n(H), assuming that atomic Hydrogen has half the SiO collisional cross section of molecular hydrogen and that ionized Hydrogen does not collide with SiO), gas temperature, velocity, molecular abundances and unresolved turbulent velocity are specified. These points are then smoothed by Lloyd's algorithm \citep{Lloyd1982} in order to minimise the variation in distance between points whilst keeping the same underlying distribution. These points are then connected by Delaunay triangulation and it is between the points connected by this method that photon are allowed to propagate (fig.~\ref{grid}). The level populations of the selected molecules are calculated at each of these points from collisional and radiative (de)excitation and the local radiation field is calculated. This is repeated 20 times with the populations of each level converging towards a single value. This number of iterations is sufficient for the signal to noise ratio of the level populations (as defined in \citealt{Brinch:2010p13078}) to exceed 1000 for 99\% of the points, ensuring that the simulation has converged on a stable level population. After 20 iterations the model is ray-traced in order to produce synthetic brightness maps. The average of ten separate runs was taken to minimise the artefacts in the output images, resulting from the grid construction.% (Fig.~\ref{points}).

@@

\subsection{Grid construction} \label{subsec:gridding}
In order to construct the grid, candidate points are randomly selected from the volume to be simulated. These candidates then have their equivalent H$_2$ number density, and the number density of SiO, compared against those of a reference point in order to decide if the candidate point is to be used in the grid or not. Candidate grid points are selected at random in a cylindrical coordinate system that is linearly spaced in z and $\theta$ and logarithmically spaced in r. For each point to be selected, a random number $\alpha$ is drawn from the semi-open set [0,$\,$1) as a threshold. After selection of random coordinates, the H$_2$ density and SiO density at the candidate point (n and m, respectively) are compared against the densities of a reference point in the un-perturbed ambient medium multiplied by a factor of $\frac{4\eta}{5}$ (n$_{ref}$ and m$_{ref}$), where $\eta$ is the density contrast between the jet and ambient medium. If $\alpha<\left( \frac{n}{n_0} \right)^{0.3}$ or $\alpha< \left( \frac{m}{m_0} \right)^{0.3}$ then the point is selected for use. Otherwise another r, $\theta$, z co-ordinate is selected and it becomes the candidate point. In addition to this method of selection, 5\% of the points are linearly distributed in x, y and z with no bias with regards to density or abundance. This provides a minimum level of sampling for the large low density regions in the outer parts of the simulated volume. %See fig. \ref{points} for an example of the points distribution in r, z. The function comparing the candidate point to the reference point and the candidate point distribution were selected empirically to sample all scales while ensuring that the majority of points are located in the inner disc where the density is higher.
\smallskip
